---
title: "Analysis"
author: "Yuka, Adela, Jack"
date: "`r Sys.Date()`"
output: html_document
---
## loading library
```{r}
library(tidyverse)
library(broom)
library(GGally)
library(lbutils)
library(olsrr)
```

```{r}
NBA <- read_csv("data/NBA.csv")
colnames(NBA)
```


```{r}
# mp, efg%, orb, drb, pf snd age
# FG% 3P, 3P%, 2P, 2P%, FT%, TRB, AST, STL, BLK, TOV, PTS？
salarymodel <- NBA %>% 
  select(-rk, -player, -player_id, -`17_18salary`, -name, -tm, -`fg`, -`fga`, -`3pa`, -`2pa`, -pos, -ft, -fta, -g, -gs, -`efg%`, -mp, -orb, - drb, -pf, -age)
glimpse(salarymodel)
```

Data exploratory analysis
```{r}
salarymodel %>%
    ggpairs()
```

```{r}
lmmodel <- lm(`16_17_salary` ~ ., data = salarymodel)
mult_sal <- tidy(lmmodel)
mult_sal
```

```{r}
g_mult_sal <- broom::glance(lmmodel)
g_mult_sal
```

```{r}
full_anova <- lb_anovat_lm(lmmodel, reg_collapse = TRUE)
full_anova

```


```{r}
qplot(data = salarymodel, x = `16_17_salary`, geom = "histogram", main = "Distribution of 2016-2017 Salary", bins = 20, color = I("black"), fill = I("blue"))
```

```{r}
qplot(data = salarymodel, y = `16_17_salary`, geom = "boxplot", main = "Distribution of 2016-2017 Salary")

```

```{r}
salarymodel %>%
  mutate(lnsalary = log(`16_17_salary`)) -> salarymodel2

qplot(data = salarymodel2, x = lnsalary, geom = "histogram", bins = 20, color = I("black"), fill = I("blue"), main = "Distribution of ln(2016-2017 Salary)")
```
```{r}
qplot(data = salarymodel2, y = lnsalary, geom = "boxplot", main = "Distribution of ln(salary)")

```

```{r}
ggplot(data = salarymodel2, aes(sample = lnsalary)) +
  stat_qq() +
  stat_qq_line(color = I("blue"))
```

```{r}
salarymodel2 %>% 
  select(-`16_17_salary`) -> salarymodel2

newmodel <- lm(lnsalary ~ ., data = salarymodel2)
newmodel_t <- tidy(newmodel)
newmodel_t
```
 
  * Fitted model:  $\widehat{salary} = 13.02 -1.023 \cdot trans.team + 5.82 \cdot fg \% + 0.0094 \cdot 3p + 0.25 \cdot 3p \% + 0.0034 \cdot 2p - 4.18 \cdot 2p \% + 0.394 \cdot ft \% + 0.0023 \cdot trb +0.0015 \cdot ast + 0.00037 \cdot stl -0.0043 \cdot blk + -0.0025 \cdot tov - 0.00078 \cdot pts$
```{r}
newmodel_g <- broom::glance(newmodel)
newmodel_g
```

```{r}
newmodel_anova <- lb_anovat_lm(newmodel, reg_collapse = TRUE)
newmodel_anova

```

```{r}
tidy(newmodel, conf.int = "TRUE", conf.level = 0.98)

```
### Method 1: All Subset Models

```{r}
all_subsets_model <- ols_step_all_possible(newmodel)
all_subsets_model
```

```{r}
plot(all_subsets_model)
```

### Method 2: The Best Subset Selection Method
```{r}
best_subset <- ols_step_best_subset(newmodel)
best_subset
```

```{r}
plot(best_subset)

```

## Stepwise Regression Techniques to Select the Best Model

### Forward Selection Procedures

```{r}
step_for_model_def <- ols_step_forward_p(newmodel, details = TRUE, progress = TRUE)  #Default p-value 

```

```{r}
step_for_model_def <- ols_step_forward_p(newmodel, details = TRUE, progress = TRUE)  #Default p-value 

```

```{r}
step_for_model_def

```

```{r}
plot(step_for_model_def)

```


```{r}
step_for_model_p15 <- ols_step_forward_p(newmodel, pent = 0.15)
step_for_model_p15
```
```{r}
plot(step_for_model_p15)

```

### backward

```{r}
step_back_mod <- ols_step_backward_p(newmodel)
step_back_mod
```

```{r}
plot(step_back_mod)

```

## step both


### step both
```{r}

step_model <- ols_step_both_p(newmodel)
step_model
```

```{r}
plot(step_model)

```
```



## When we want to do prediction
```{r}
# newobs <- data.frame(bedrms = 4, baths = 2, size = 2000, yard = 7500, age = 45, nbhd = 2)
# predict(multmodprice, newdata = newobs, interval = "confidence", level = 0.95)

# newobs <- data.frame(bedrms = 4, baths = 2, size = 2000, yard = 7500, age = 45, nbhd = 2)
# predict(multmodprice, newdata = newobs, interval = "prediction", level = 0.95)
```

