---
title: "Analysis"
author: "Yuka, Adela, Jack"
date: "`r Sys.Date()`"
output: html_document
---
## loading library
```{r}
library(tidyverse)
library(broom)
library(GGally)
library(lbutils)
library(olsrr)
library(fastDummies)
library(lattice)
library(ggfortify)
library(car)

```

```{r}
NBA <- read_csv("NBA.csv")
colnames(NBA)
```

```{r}
salarymodel <- NBA %>% 
  select(-rk, -player, -player_id, -`17_18salary`, -name, -tm, -`fg`, -`fga`, -`3pa`, -`2pa`, -ft, -fta, -g, -gs, -`efg%`, -mp, -orb, - drb, -pf, -age)

glimpse(salarymodel)
```

Data exploratory analysis
```{r}
salarymodel %>%
    ggpairs()
```

```{r}
lmmodel <- lm(`16_17_salary` ~ ., data = salarymodel)
mult_sal <- tidy(lmmodel)
mult_sal
```

```{r}
g_mult_sal <- broom::glance(lmmodel)
g_mult_sal
```

```{r}
full_anova <- lb_anovat_lm(lmmodel, reg_collapse = TRUE)
full_anova

```


```{r}
qplot(data = salarymodel, x = `16_17_salary`, geom = "histogram", main = "Distribution of 2016-2017 Salary", bins = 20, color = I("black"), fill = I("blue"))
```

```{r}
qplot(data = salarymodel, y = `16_17_salary`, geom = "boxplot", main = "Distribution of 2016-2017 Salary")

```

```{r}
salarymodel %>%
  mutate(lnsalary = log(`16_17_salary`)) -> salarymodel
```


```{r}
qplot(data = salarymodel2, x = lnsalary, geom = "histogram", bins = 20, color = I("black"), fill = I("blue"), main = "Distribution of ln(2016-2017 Salary)")
```

```{r}
vif(salarymodel2)
```

```{r}
qplot(data = salarymodel2, y = lnsalary, geom = "boxplot", main = "Distribution of ln(salary)")

```

```{r}
ggplot(data = salarymodel2, aes(sample = lnsalary)) +
  stat_qq() +
  stat_qq_line(color = I("blue"))
```

## full model
```{r}
salarymodel %>% 
  select(-`16_17_salary`, -pos , -trans_team) -> salarymodel2

newmodel <- lm(lnsalary ~ ., data = salarymodel2)
newmodel_t <- tidy(newmodel) 
newmodel_t 
```
```{r}
vif(newmodel)
```
 
  * Fitted model:  $\widehat{salary} = 13.02 -1.023 \cdot trans.team + 5.82 \cdot fg \% + 0.0094 \cdot 3p + 0.25 \cdot 3p \% + 0.0034 \cdot 2p - 4.18 \cdot 2p \% + 0.394 \cdot ft \% + 0.0023 \cdot trb +0.0015 \cdot ast + 0.00037 \cdot stl -0.0043 \cdot blk + -0.0025 \cdot tov - 0.00078 \cdot pts$
  
```{r}
newmodel_g <- broom::glance(newmodel)
newmodel_g
```

```{r}
newmodel_anova <- lb_anovat_lm(newmodel, reg_collapse = FALSE)
newmodel_anova

```

```{r}
tidy(newmodel, conf.int = "TRUE", conf.level = 0.98)

```
## Reuced Model based on subset thing
```{r}

reduce <- NBA %>% 
  mutate(lnsalary = log(`16_17_salary`)) %>% 
  select(lnsalary,`fg%`,`3p`,`3p%`,`2p%`,trb,blk, pos, trans_team)

reducemodel <- lm(lnsalary ~`fg%` + `3p` + `3p%`+ `2p%`+ trb + blk, data = salarymodel2)

reduce_model_t <- tidy(reducemodel) 
reduce_model_t

reduce_model_g <- broom::glance(reducemodel)
reduce_model_g

reduce_model_a <- lb_anovat_lm(reducemodel, reg_collapse = FALSE)
reduce_model_a

tidy(reducemodel, conf.int = "TRUE", conf.level = 0.98)

vif(reducemodel)


```


```{r}
# dummy vaiable
salarymodel %>%
  mutate(lnsalary = log(`16_17_salary`)) %>% 
  select(-'16_17_salary')-> 
  dummy
```


```{r}
results <- dummy_cols(.data = reduce, select_columns = c("pos","trans_team"))

results %>%
  select(pos, pos_C, pos_PF, pos_PG, pos_SF, pos_SG, trans_team, trans_team_none, trans_team_trans) %>%
  head(6)

newresult <- dummy_cols(.data = reduce, select_columns = c("pos","trans_team"), remove_selected_columns = TRUE)

rename(.data = newresult, trans = trans_team_trans) -> newdummy

dummy_model <- lm(lnsalary ~ ., data = newdummy)

dumtidyout <- tidy(dummy_model)
dumglout <- glance(dummy_model)
dumtidyout
dumglout
```

# Comparing models

```{r}
dumglout #dummy models
```

```{r}
newmodel_g #full model wihtout dummy
```

* For the full model including the dummy variables $R^2_{adj}$ = 0.4492.

* For the full model without the dummy variables $R^2_{adj}$ = 0.4096

### Method 1: All Subset Models

```{r}
all_subsets_model <- ols_step_all_possible(newmodel)
all_subsets_model

all_subsets_model %>% 
  mutate(bestcp = cp - n) %>% 
  arrange(bestcp)-> x

```

```{r}
plot(all_subsets_model)
```

### Method 2: The Best Subset Selection Method
```{r}
best_subset <- ols_step_best_subset(newmodel)
best_subset
```

```{r}
plot(best_subset)

```

## Stepwise Regression Techniques to Select the Best Model

### Forward Selection Procedures

```{r}
step_for_model_def <- ols_step_forward_p(newmodel, details = TRUE, progress = TRUE)  #Default p-value 

```

```{r}
step_for_model_def <- ols_step_forward_p(newmodel, details = TRUE, progress = TRUE)  #Default p-value 

```

```{r}
step_for_model_def

```

```{r}
plot(step_for_model_def)

```


```{r}
step_for_model_p15 <- ols_step_forward_p(newmodel, pent = 0.15)
step_for_model_p15
```
```{r}
plot(step_for_model_p15)

```

### backward

```{r}
step_back_mod <- ols_step_backward_p(newmodel)
step_back_mod
```

```{r}
plot(step_back_mod)

```

## step both
### step both
```{r}

step_model <- ols_step_both_p(newmodel)
step_model
```

```{r}
plot(step_model)

```




### When we want to do prediction

```{r}
test <- read_csv("player17_18.csv")
names(test)[1:31] <- tolower(names(test)[1:31])

test_result <- dummy_cols(.data = test, select_columns = c("pos","trans_team"), remove_selected_columns = TRUE)
rename(.data = test_result, trans = trans_team_trans) -> test_dummy

# final reduce model
full_predict <- predict(dummy_model, newdata = test_dummy, interval = "confidence", level = 0.95)
full_predict <- cbind(test_dummy, full_predict)

full_predict1 <- full_predict %>% 
  left_join(NBA, by = c("player_id" = "player_id", "tm" = "tm")) %>% 
  select(fit, `17_18salary`)
  
diff <- log(full_predict1$`17_18salary`)-full_predict1$fit

MAD <- mean(abs(diff),na.rm = TRUE)
MSE <- mean(diff^2,na.rm = TRUE)



```

# model testing
```{r}
# Assumtion
autoplot(dummy_model)

# Residual normality test
shapiro.test(dummy_model$residuals)

# Residual independence test
durbinWatsonTest(dummy_model)

# Residual variance homogeneity test
ncvTest(dummy_model)
```

```{r}

refit <- lm(abs(residuals(dummy_model))~fitted(dummy_model))
refit

wts <- 1 / fitted(refit)^2
wts

lm_wls <- lm(lnsalary ~ ., data = newdummy, weights = wts)
summary(lm_wls)

plot(refit)
autoplot(lm_wls)
ncvTest(lm_wls)


# Residual normality test
# shapiro.test(lm_wls$residuals)

# Residual independence test
durbinWatsonTest(lm_wls)


```

