---
title: "Final_Project_Clean"
author: "Yuka, Adela, Jack"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(broom)
library(GGally)
library(lbutils)
library(olsrr)
library(fastDummies)
library(lattice)
library(ggfortify)
library(car)
```

#### Loading Data 
```{r}
NBA <- read_csv("NBA.csv")
glimpse(NBA)
```

#### Original Model - lnsalary with 15 variables


```{r}
original_data <- NBA %>% 
  select(-rk, -player, -player_id, -`17_18salary`, -name, -tm, -`fg`, -`fga`, -`3pa`, -`2pa`, -ft, -fta, -g, -gs, -`efg%`, -mp, -orb, - drb, -pf, -age)

glimpse(original_data)
```

### Corralation Analysis
```{r}
original_data %>%
    ggpairs()
```

## Checking Y Preditable 
```{r}
qplot(data = original_data, x = `16_17_salary`, geom = "histogram", main = "Distribution of 2016-2017 Salary", bins = 20, color = I("black"), fill = I("blue"))
```

```{r}
qplot(data = original_data, y = `16_17_salary`, geom = "boxplot", main = "Distribution of 2016-2017 Salary")

```
## Log Salary for Better Prediction 
```{r}
original_data %>%
  mutate(lnsalary = log(`16_17_salary`)) -> original_data
```
 
```{r}
qplot(data = original_data, x = lnsalary, geom = "histogram", bins = 20, color = I("black"), fill = I("blue"), main = "Distribution of ln(2016-2017 Salary)")
```

## Original Model

```{r}

original_data %>% 
  select(-`16_17_salary`, -pos , -trans_team) -> new_og_data

og_model <- lm(lnsalary ~ ., data = new_og_data)

mult_og <- tidy(og_model)
mult_og
```

 * Fitted model:  $\widehat{salary} = 13.02 -1.023 \cdot trans.team + 5.82 \cdot fg \% + 0.0094 \cdot 3p + 0.25 \cdot 3p \% + 0.0034 \cdot 2p - 4.18 \cdot 2p \% + 0.394 \cdot ft \% + 0.0023 \cdot trb +0.0015 \cdot ast + 0.00037 \cdot stl -0.0043 \cdot blk + -0.0025 \cdot tov - 0.00078 \cdot pts$
 
```{r}
g_mult_og <- broom::glance(og_model)
g_mult_og
```

```{r}
og_anova <- lb_anovat_lm(og_model, reg_collapse = TRUE)
og_anova
```

```{r}
vif(og_model)
## shows our original model isn't good enought to be our final model
```

### All Subset Models

```{r}
all_subsets_model <- ols_step_all_possible(og_model)
all_subsets_model
```

```{r}
plot(all_subsets_model)
```

### Reduce Model based on All Subset Models Method
```{r}
reduce <- NBA %>% 
  mutate(lnsalary = log(`16_17_salary`)) %>% 
  select(lnsalary,`fg%`,`3p`, `2p`, `2p%`,trb, pos, trans_team)

```

```{r}
reducemodel <- lm(lnsalary ~`fg%` + `3p` + `2p` + `2p%`+ trb, data = new_og_data)

reduce_model_t <- tidy(reducemodel) 
reduce_model_t
```


```{r}
reduce_model_g <- broom::glance(reducemodel)
reduce_model_g
```


```{r}
reduce_model_a <- lb_anovat_lm(reducemodel, reg_collapse = FALSE)
reduce_model_a
```


```{r}
tidy(reducemodel, conf.int = "TRUE", conf.level = 0.98)
```


```{r}
vif(reducemodel)
```

## Plotting Reduce Model 
```{r}
# Assumtion
autoplot(reducemodel)

# Residual normality test
shapiro.test(reducemodel$residuals)

# Residual independence test
durbinWatsonTest(reducemodel)

# Residual variance homogeneity test
ncvTest(reducemodel)
```


## Adding Dummy Variables
```{r}
results <- dummy_cols(.data = reduce, select_columns = c("pos","trans_team"))

results %>%
  select(pos, pos_C, pos_PF, pos_PG, pos_SF, pos_SG, trans_team, trans_team_none, trans_team_trans) %>%
  head(6)

newresult <- dummy_cols(.data = reduce, select_columns = c("pos","trans_team"), remove_selected_columns = TRUE)

rename(.data = newresult, trans = trans_team_trans) -> newdummy

dummy_model <- lm(lnsalary ~ ., data = newdummy)
```


```{r}
dumtidyout <- tidy(dummy_model)

dumglout <- glance(dummy_model)

dumtidyout
dumglout

dummy_a <- lb_anovat_lm(dummy_model, reg_collapse = FALSE)
dummy_a
```
#### Comparing models

```{r}
dumglout #dummy models
```

```{r}
g_mult_og #full model wihtout dummy
```

```{r}
reduce_model_g # reduce model
```


#### Model Testing

```{r}
library(lmtest)
bptest(dummy_model)
```

```{r}
# Assumtion
autoplot(dummy_model)

# Residual normality test
shapiro.test(dummy_model$residuals)

# Residual independence test
durbinWatsonTest(dummy_model)

# Residual variance homogeneity test
ncvTest(dummy_model)
```

###Fixing Heteroskedasticity by Using WLS

```{r}
refit <- lm(abs(residuals(dummy_model)) ~ fitted(dummy_model))
refit
```


```{r}
wts <- 1 / fitted(refit)^2
wts
```

### Final Model
```{r}
lm_wls <- lm(lnsalary ~ ., data = newdummy, weights = wts)

tidy(lm_wls)

```


```{r}
glance(lm_wls)
```

```{r}
final_model_a <- lb_anovat_lm(lm_wls)
final_model_a
```


```{r}
# Assumtion
autoplot(lm_wls)

# Residual normality test
shapiro.test(lm_wls$residuals)

# Residual independence test
durbinWatsonTest(lm_wls)

# Residual variance homogeneity test
ncvTest(lm_wls)

```

### Evaluate Forecase Model

```{r}
test <- read_csv("player17_18.csv")
names(test)[1:31] <- tolower(names(test)[1:31])

test_result <- dummy_cols(.data = test, select_columns = c("pos","trans_team"), remove_selected_columns = TRUE)
rename(.data = test_result, trans = trans_team_trans) -> test_dummy

# final reduce model
full_predict <- predict(lm_wls, newdata = test_dummy, interval = "confidence", level = 0.95)
full_predict <- cbind(test_dummy, full_predict)

full_predict1 <- full_predict %>% 
  left_join(NBA, by = c("player_id" = "player_id", "tm" = "tm")) %>% 
  select(fit, `17_18salary`)
  
diff <- log(full_predict1$`17_18salary`)-full_predict1$fit

MAD <- mean(abs(diff),na.rm = TRUE)
MSE <- mean(diff^2,na.rm = TRUE)

```